---
title: "Classification"
output: html_notebook
---

# {.tabset}

## Overview

First, I will try rebuilding the classifier present in Heidi's "inventory" repository.

```{r lsetup, echo =FALSE}
library(dplyr)
library(stringr)
```


## Initial Classifier

First, import the manually classified set:

```{r read_data}
manual_checks <- read.csv("../data/processed/manual_checks_hji_2021-08-20.csv")
```

Next run the classification:

`dplyr::case_when()` goes through the various conditions. Upon the first `TRUE` evaluation, the value is assigned. So earlier cases take precence.

*Note:* Currently `db` is being classified as "db" not as database.

```{r}
manual_checks <- manual_checks %>%
  mutate(
    title = tolower(title),
    title_cat = factor(title_cat),
    title_class = case_when(
      str_detect(title, "database") ~ "database",
      str_detect(title, "data") ~ "data",
      str_detect(title, "base\\b") ~ "base",
      str_detect(title, "db") ~ "db",
      str_detect(title, "bank") ~ "bank",
      str_detect(title, "package") ~ "package",
      str_detect(title, "tool") ~ "tool",
      str_detect(title, "server") ~ "server",
      str_detect(title, "study") ~ "study",
      TRUE ~ "other"
    ))
```


How many titles were given each classification?

```{r total_classified}
manual_checks %>%
  group_by(title_class) %>% 
  count()
```

Based on true classification, how many were correctly classified?

* `count`: Number of observations given that classification manually
* `n_correct`: Number observations correctly classified
* `prop_correct`: Proportion of that category correctly classified

```{r class_results}
class_summary <- manual_checks %>% 
  group_by(title_cat) %>% 
  summarize(count = n()) %>% 
  complete(title_cat, fill = list(count = 0))

class_summary <- class_summary %>%
  mutate(
    manual_checks %>%
      group_by(title_cat) %>%
      mutate(correct_class = (title_cat == title_class)) %>%
      filter(correct_class == TRUE) %>%
      summarize(n_correct = n()) %>%
      complete(title_cat, fill = list(n_correct = 0)) %>%
      select(n_correct),
    prop_correct = round(n_correct / count, 3)
  )

class_summary
```

Many of the classifications are pretty good. Notice though that there is not `database` included, because none of the manually checked items have that category

```{r}
any(manual_checks$title_cat == "database")
```

